---
title: "Class 08: Mini Project Breast Cancer"
author: "Anthony Truong A16906178"
format: pdf
toc: true
---

## Background

In today's class we will be employing all the R techniques for data analysis that we have learned thus far - including the machine learning methods of clustering and PCA - to analyze real breast cancer biopsy data. 

## Data Import 

The data is in an CSV format:

```{r}
wisc.df <- read.csv("WisconsinCancer.csv", row.names = 1)
```
Small peek at the data 
```{r}
head(wisc.df, 3)
```

> Q1 How many observations are in this dataset? 

```{r}
nrow(wisc.df)
```


> Q2 How many observations have a malignant diagnosis?


```{r}
table(wisc.df$diagnosis == "M")
```

>Q3 How many variables/features in the data are suffixed with _mean? 

```{r}
length(grep("_mean", colnames(wisc.df)))
```

We need to remove the `diagnosis` column before we do any further analysis of this dataset - we don't want to pass this to PCA etc. We will save it as a separate small vector that we can use later to compare our findings to those of experts. 

```{r}
wisc.data <- wisc.df[,-1]
diagnosis <- wisc.df$diagnosis
```

## Principal Component Analysis (PCA)

The main function in base R is called `prcomp()`we will use the optional argument `scale = TRUE` here as the data columns /features/dimensions are on very different scales in the original data set. 

```{r}
wisc.pr <- prcomp(wisc.data, scale = TRUE)
```

```{r}
attributes(wisc.pr)
```

```{r}
library(ggplot2)

ggplot(wisc.pr$x) + 
  aes (PC1, PC2, col = diagnosis) + 
  geom_point()
```

```{r}
colMeans(wisc.data)
apply(wisc.data, 2, sd)
```

```{r}
summary(wisc.pr)
```


> Q4. From your results, what proportion of the original variance is captured by the first principal component (PC1)?

```{r}
var_explained <- wisc.pr$sdev^2 / sum(wisc.pr$sdev^2)
var_explained[1] 
```



> Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

```{r}
var_explained <- wisc.pr$sdev^2 / sum(wisc.pr$sdev^2)
cumulative <- cumsum(var_explained)

which(cumulative >= 0.70)[1]
```


> Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

```{r}
var_explained <- wisc.pr$sdev^2 / sum(wisc.pr$sdev^2)
cumulative <- cumsum(var_explained)

which(cumulative >= 0.90)[1]
```
> Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?

```{r}
biplot(wisc.pr)
```


The plot seems to stand out a lot and has a big area of cluters in the middle with numerous arrows pointing towards the dimensions plotted on the graph.

> Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots?

```{r}
ggplot(wisc.pr$x) +
  aes(PC1, PC3, col=diagnosis) +
  geom_point()
```

The plots on this graph have the points much lower but in similar clusters compared to the last graph. 

> Q9. For the first principal component, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature concave.points_mean? This tells us how much this original feature contributes to the first PC. Are there any features with larger contributions than this one?

The component of the loading vector for the concave points means finding the original measurements that contribute the most to the `wisc.pr` data set in the first column. That is then translated into the concave points mean function. There are not any functions that are higher than the concave.points_mean. 

```{r}
wisc.pr$rotation [ ,1]
```

# 4. Hierarchial Clustering 

The goal of this section is to do hierarchical clustering of the original data to see if there is any obvious grouping into malignant and benign clusters. 

```{r}
data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled, method = "euclidian")
wisc.hclust <- hclust(data.dist, method = "complete")
```


```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k = 2)
table(wisc.hclust.clusters)
```
> Q10. Using the plot() and abline() functions, what is the height at which the clustering model has 4 clusters?

```{r}
plot(wisc.hclust)
abline(wisc.hclust, col="red", lty=2)
```
The height where there is 4 clusters is at 15. 

> Q12. Which method gives your favorite results for the same data.dist dataset? Explain your reasoning.

The method that gives me my favorite results from the dataset is the latest one we just did. As it gives us more specific results as well as keeping it accurate and identifying the variance in the data-set. 

## Combining Methods 

The idea here is that I can take my new variables (the PCs) that are better descriptors of the data-set than the original features (i.e. the 30 columns in `wisc.data`) and use these as a basis for clustering. 

```{r}
pc.dist <- dist(wisc.pr$x[ ,1:3])
wisc.pr.hclust <- hclust (pc.dist, method = "ward.D2")
plot(wisc.pr.hclust)
```

```{r}
grps <- cutree(wisc.pr.hclust, k = 2)
table(grps)
```

I can now run `table()` with both my clustering `grps` and the expert `diagonosis` 

```{r}
table(grps, diagnosis)
```
> Q13.  How well does the newly created hclust model with two clusters separate out the two “M” and “B” diagnoses?

The new `hclust` model breaks down the groups into more selective groups of M and B allowing for a more closed dendrite graph shown above. It is more visually appealing as shown above. 

Our cluster "1" has 179 "M" diagnosis 
Our cluster "2" has 333 "B" diagnosis 

179 TP 
24 FP 
333 TN 
24 FN 

Sensitivity: TP/(TP + FN)
```{r}
179/(179+33)
```
Specificity: TN/(TN +FP)
```{r}
333/(333+24)
```

## Prediction 

We can use our PCA model for prediction of new un-seen cases.

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```

> Q16. Which of these new patients should we prioritize for follow up based on your results?

Based on the results we should prioritze patient 1 as the clusters have a more specific and original points displayed on the graph in comparison to patient 2 which is more sporadic. 