---
title: "Class 17: Downstream + PCA Analysis"
author: "Anthony Truong (A16906178)"
format: pdf
toc: TRUE
---

## Installing txt import package 

```{r}
# BiocManager::install("tximport")
```

```{r}
library(tximport)

# setup the folder and file-names to read
folders <- dir(pattern="SRR21568*")
samples <- sub("_quant", "", folders)
files <- file.path( folders, "abundance.h5" )
names(files) <- samples

txi.kallisto <- tximport(files, type = "kallisto", txOut = TRUE)
```


```{r}
head(txi.kallisto$counts)
```

Transcript count: 

```{r}
colSums(txi.kallisto$counts)
```

How many transcripts are detected in a single sample! 
```{r}
sum(rowSums(txi.kallisto$counts)>0)
```

It needs to be filtered out before subsequent analysis: 

```{r}
to.keep <- rowSums(txi.kallisto$counts) > 0
kset.nonzero <- txi.kallisto$counts[to.keep,]
```
Those with no change: 
```{r}
keep2 <- apply(kset.nonzero,1,sd)>0
x <- kset.nonzero[keep2,]
```

## PCA Analysis 

We can now apply any exploratory analysis technique to this counts matrix. As an example, we will perform a PCA of the transcriptomic profiles of these samples.

Now we compute the principal components, centering and scaling each transcript’s measured levels so that each feature contributes equally to the PCA:

```{r}
pca <- prcomp(t(x), scale=TRUE)
summary(pca)
```

This can be summarized on an plot: 

```{r}
plot(pca$x[,1], pca$x[,2],
     col=c("blue","blue","red","red"),
     xlab="PC1", ylab="PC2", pch=16)
```

> Use ggplot to make a similar figure of PC1 vs PC2 + separate figures for PC1 vs PC3 and PC2 vs PC3? 

```{r}
library(ggplot2)
library(ggrepel)

mycols <- c("blue","blue","red","red")

ggplot(pca$x) +
  aes(PC1, PC2, label=rownames(pca$x)) +
  geom_point( col=mycols ) +
  geom_text_repel( col=mycols ) +
  theme_bw()
```


```{r}
library(ggplot2)
library(ggrepel)

mycols <- c("blue","blue","red","red")

ggplot(pca$x) +
  aes(PC1, PC3, label=rownames(pca$x)) +
  geom_point( col=mycols ) +
  geom_text_repel( col=mycols ) +
  theme_bw()
```


```{r}
library(ggplot2)
library(ggrepel)

mycols <- c("blue","blue","red","red")

ggplot(pca$x) +
  aes(PC2, PC3, label=rownames(pca$x)) +
  geom_point( col=mycols ) +
  geom_text_repel( col=mycols ) +
  theme_bw()
```
```

