---
title: "Class012: Lab Session"
author: "Anthony Truong (A16906178)"
format: pdf
toc: true
---

## Section 1: Proportion of G/G in a population 

Downloaded the CSV file from ENSEMBL <https://useast.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39894595-39895595;v=rs8067378;vdb=variation;vf=959672880#373531_tablePanel>

Here we read this CSV file 

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl) * 100
```

Now let's look at a different population. I picked the GBR population. 

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378(1).csv")
head(gbr)
```

Find proportion of G|G 

```{r}
round(table(gbr$Genotype..forward.strand.) / nrow(gbr) * 100, 2)
```

This variant that is associated with childhood asthma is more frequent in the GBR population than the MKL population. 

Let's now dig into this further! 

## Section 4: Population Scale Analysis 

One sample is obviously not enough to know what is happening in a population. You are interested in assessing genetic differences on a population scale. 

How many samples do we have? 

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

```{r}
nrow(expr)
```

```{r}
table(expr$geno) / nrow(expr) * 100
```

```{r}
library(ggplot2)
```

Let's make an box plot! 

```{r}
ggplot(expr) + aes(geno, exp, fill = geno) +
  geom_boxplot(notch = TRUE)
```

It does seem SNP effects the expression of ORMDL3 as the box plot on the graph seems to go down on corresponding geno-types being matched with each other. A|A is higher in ORMDL3 than G|G. A|G is in the middle of the pack between these two values. 