---
title: "Class06 R Functions"
author: "Anthony Truong (A16906178)"
format: pdf
toc: true
---

## Background 

Functions are at the heart of using R. Everything we do involves calling and using functions (from data input, analysis to results output). 

All functions in R have at least 3 things:

1. A **name** the thing we use to call the function. 
2. One or more input **arguments** that are comma separated.
3. The **body**, lines of code between curly brackets { } that does the work of the function. 

## A first function 

Let's write a silly function to add some numbers to: 

```{r}
add <- function(x) {
  x + 1 
}
```

Let's try it out 
```{r}
add(100)
```

Will this work?

```{r}
add(c(100, 200, 300) )
```
Modify this to be more useful and add more than just 1 

```{r}
add <- function(x, y = 1) {
  x + y 
}
```


```{r}
add(100,10)
```
Will this work? 
```{r}
add(100)
```

```{r}
plot(1:10, col = "blue", typ = "b")
```

```{r}
log(10, base = 10)
```

> **N.B.** Input arguments can either be **required** or an **optional**. The later have a fall-back default that is specified in the function code with an equals sign. 


```{r}
#add(x = 100, y = 200, z = 300)
```


## A second function

All functions in R look like this 

```
name <- function(arg){
body 
}

```
The `sample()` function in R randomly selects items from a vector. 

```{r}
sample(1:10, size = 4)
```

> Q. Return 12 numbers picked randomly from the input 1:10 

```{r}
sample(1:10, size = 12, replace = TRUE)
```

> Q. Write the code to generate a random 12 nucleotide long DNA sequence? 

```{r}
bases <- c("A","C","G","T")
sample(bases, size = 12, replace = TRUE)
```

> Q. Write a first version function called `generate_dna()` that generates a user specified length `n` random DNA sequence?


```
name <- function(arg){
body 
}

```

```{r}
generate_dna <- function(n = 6){
 bases <- c("A","C","G","T") 
 sample(bases, size = n , replace = TRUE)
}
generate_dna()
```

> Q. Modify your function to return a FAST like sequence so rather than [1] "T" "T" "G" "A" "A" "T" we want "TTGAAT" 


```{r}
generate_dna <- function(n = 6){
 bases <- c("A","C","G","T")
 ans <- sample(bases, size = n, replace = TRUE)
 ans <- paste(ans, collapse = "")
 return(ans)
 x <- "p"
 x
}
```

```{r}
generate_dna()
```

> Q. Give the user an option to return FASTA format output sequence or standard multi-element vector format?

```{r}
generate_dna <- function(n = 6, fasta = TRUE) {
 bases <- c("A","C","G","T")
 ans <- sample(bases, size = n, replace = TRUE)
 
 if(fasta) { 
 ans <- paste(ans, collapse = "")
 cat("Hello...")
 } else{
cat("...is it me you're looking for...")   
 }
 
 return(ans)
}
```

```{r}
generate_dna(10)
generate_dna(10, fasta = FALSE)
```

## A new cool function 

> Q. Write a function called `generate_protein()` that generates a user specified length protein sequence in FASTA like format?

```{r}
generate_protein <- function(n = 4) {
  aa <- c("A","R","N","D","C","E","Q","G","H","I",
  "L","K","M","F","P","S","T","W","Y","V")
 ans <- sample(aa, size = n, replace =TRUE)
 ans <- paste(ans, collapse = "")
return(ans)
}
```

```{r}
generate_protein()
```

> Q. Use your new `generate_protein()`function to generate all sequences between length 6 and 12 amino -acids in length and check if any of these are unique in nature (i.e. found in the NR database at NCBI)?

```{r}
generate_protein(6)
generate_protein(7)
generate_protein(8)
generate_protein(9)
generate_protein(10)
generate_protein(11)
generate_protein(12)
```

Blasting these, it seems 6/7/10 aren't unique proteins and don't have any unique identifiers to them. But on the 8/9/11/12 line protein its not seen in nature and are unique. 

Or we could for a `for()` loop: 

```{r}
for(i in 6:12) {
  cat(sep ="" ,">", i, "\n")
  cat(generate_protein(i), "\n")
}
```


